### 2017-11-15 doc/ms.tex
include Noebauer figure in paper and add description of individual models

%%
%% Beginning of file 'sample61.tex'
%%
%% Modified 2016 September
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.1 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
%% All of these external packages should already be present in the modern TeX 
%% distributions.  If not they can also be obtained at www.ctan.org.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.1
\documentclass[twocolumn]{aastex61}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be envoked like this:
%%
%% \documentclass[argument]{aastex61}
%% 
%% where the arguement options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can envoke to allow other stylistic
%% actions. The available options are:
%%
%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%  tighten      : Makes baselineskip slightly smaller, only works with 
%%                 the twocolumn substyle.
%%  times        : uses times font instead of the default
%%  linenumbers  : turn on lineno package.
%%  trackchanges : required to see the revision mark up and print its output
%%  longauthor   : Do not use the more compressed footnote style (default) for 
%%                 the author/collaboration/affiliations. Instead print all
%%                 affiliation information after each name. Creates a much
%%                 long author list but may be desirable for short author papers
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex61}

%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003

%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\newcommand{\sm}{M_\odot}
\newcommand{\sr}{R_\odot}
\newcommand{\abc}{iPTF\,16abc}
\newcommand{\sneia}{SNe Ia}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{\today}
\revised{}
\accepted{}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{ApJ}

%% Mark up commands to limit the number of authors on the front page.
%% Note that in AASTeX v6.1 a \collaboration call (see below) counts as
%% an author in this case.
%
%\AuthorCollaborationLimit=3
%
%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
%% collaboration" on the front page of this example manuscript.
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{\abc}
\shortauthors{Miller et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
\watermark{DRAFT}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section defines new commands for comments from co-authors
%%
\newcommand{\ycao}[1]{{\color{red} ycao: {#1}}}
\newcommand{\amiller}[1]{{\color{blue} amiller: {#1}}}
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

\begin{document}

\title{Early Observations of \abc:\\
\textbf{a Type Ia Supernova Featuring Either} Strong Ejecta Mixing and/or Interaction with Diffuse Material}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.1 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affilations. The change was necessary to implement 
%% autoindexing of affilations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% arguement which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.1 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.1 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\correspondingauthor{A.~A.~Miller}
\email{amiller@northestern.edu}

\author[0000-0001-9515-478X]{A.~A.~Miller}
\affil{Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA) and Department of Physics and Astronomy, Northwestern University, 2145 Sheridan Road, Evanston, IL 60208, USA}
\affil{The Adler Planetarium, Chicago, IL 60605, USA}

\author[0000-0002-8036-8491]{Y.~Cao}
\affil{eScience Institute and Astronomy Department, University of Washington,
  Seattle, WA 98195}

\author[0000-0001-6806-0673]{A.~L.~Piro}
\affil{The Observatories of the Carnegie Institution for Science, 813 Santa Barbara Street, Pasadena, CA 91101, USA}

\author{N.~Blagorodnova}
\affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}

\author{B.~D.~Bue}
\affil{Jet Propulsion Laboratory, California Institute of Technology, Pasadena, CA 91109, USA}

\author[0000-0003-1673-970X]{S.~B.~Cenko}
\affil{NASA Goddard Space Flight Center, Mail Code 661, Greenbelt, MD 20771, USA}
\affil{Joint Space-Science Institute, University of Maryland, College Park, MD 20742, USA}

\author{S.~Dhawan}
\affil{The Oskar Klein Centre, Department of Physics, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}

\author{R.~Ferretti}
\affil{The Oskar Klein Centre, Department of Physics, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}

\author[0000-0003-2238-1572]{O.~D.~Fox}
\affil{Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA}

\author{C.~Fremling}
\affil{The Oskar Klein Centre, Department of Astronomy, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}

\author{A.~Goobar}
\affil{The Oskar Klein Centre, Department of Physics, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}

\author[0000-0003-4253-656X]{D.~A.~Howell}
\affil{Las Cumbres Observatory, Goleta, CA 93117, USA}
\affil{Physics Department, University of California, Santa Barbara, CA 93106, USA}

\author[0000-0002-0832-2974]{G.~Hosseinzadeh}
\affil{Las Cumbres Observatory, Goleta, CA 93117, USA}
\affil{Physics Department, University of California, Santa Barbara, CA 93106, USA}

\author[0000-0002-5619-4938]{M.~M.~Kasliwal}
\affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}

\author{R.~R.~Laher}
\affil{Infrared Processing and Analysis Center, California Institute of Technology, Pasadena, CA 91125, USA}

\author{R.~Lunnan}
\affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}

\author{F.~J.~Masci}
\affil{Infrared Processing and Analysis Center, California Institute of Technology, Pasadena, CA 91125, USA}

\author[0000-0001-5807-7893]{C.~McCully}
\affil{Las Cumbres Observatory, Goleta, CA 93117, USA}
\affil{Physics Department, University of California, Santa Barbara, CA 93106, USA}

\author[0000-0002-3389-0586]{P.~E.~Nugent}
\affil{Lawrence Berkeley National Laboratory, Berkeley, California 94720, USA}
\affil{University of California -- Berkeley, Berkeley, CA 94720, USA}

\author{J.~Sollerman}
\affil{The Oskar Klein Centre, Department of Astronomy, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}

\author{F.~Taddia}
\affil{The Oskar Klein Centre, Department of Astronomy, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}

\author[0000-0001-5390-8563]{S. R. Kulkarni}
\affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}

% \author{The Alphabet Friends}
% \affil{the intermediate Palomar Transient Factory}

%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX 
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.1 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands 
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The 
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}

Early observations of Type Ia supernovae (SNe) provide a unique probe of their
progenitor systems and explosion physics. Here, we report the intermediate
Palomar Transient Factory (iPTF) discovery of an extraordinarily young SN Ia,
\abc. By fitting a power law to our early light curve, we infer that first
light for the SN, \textbf{that is when the SN produces appreciable optical
radiation following shock breakout}, \textbf{occurred only}
$0.15\pm_{0.07}^{0.15}$\,d before our first detection. In the $\sim$24\,hr
after discovery, \abc\ rose by $\sim$2\,mag, \textbf{featuring} a near-linear
rise in flux for $\gtrsim$3 d. \textbf{Early spectra show strong \ion{C}{2}
absorption, which disappears after $\sim$7\,d.} Unlike the
extensively-observed Type Ia SN\,2011fe, the $(B-V)_0$ colors of \abc\ are
blue and nearly constant in the days after explosion. We show that our early
observations of \abc\ cannot be explained by either SN shock breakout and the
associated, subsequent cooling, or the SN ejecta colliding with a stellar
companion. Instead, we argue that the early characteristics of \abc,
including: (i) the rapid, near-linear rise, (ii) the non-evolving blue colors,
and (iii) the strong absorption from ionized carbon, are the result of either
vigorous mixing of radioactive $^{56}$Ni in the SN ejecta, or ejecta
interaction with diffuse material, or a combination of the two. In the next
few years, dozens of very young \textit{normal} SNe Ia will be discovered, and
observations similar to those presented here will constrain the white dwarf    
explosion mechanism.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{methods: observational --- surveys --- supernovae: general --- supernovae: individual (\abc)}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.

%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction}
\label{sec:intro}

Although Type Ia supernovae (SNe Ia) have been extensively used as
standardizable candles, their progenitor systems and explosion
physics are still debated (see a recent review by
\citealt{2014ARA&A..52..107M}). Extremely detailed
observations in the hours to days after explosion provide a promising avenue to further constrain this problem.

While the shock breakout of a SN Ia occurs on a sub-second timescale,
the subsequent quasi-adiabatic expansion and cooling of the unbound
ejecta produces thermal emission that can be used to infer the
radius of the exploding star
\citep{2010ApJ...708..598P,2011ApJ...728...63R}. Comparing models of this cooling emission to the earliest-phase data of SN~2011fe,
\citet{2012ApJ...744L..17B} concluded that the explosion came from a star with $R_\ast \lesssim 0.02\;\sr$, where $\sr$ is the solar
radius. Combining the radius constraint with the measured ejecta 
mass, \citeauthor{2012ApJ...744L..17B} derive the mean density of the progenitor star, confirming that at least some Type Ia SNe come from compact and degenerate stars.

Early phase observations of SNe Ia from a white dwarf (WD)$+$non-degenerate binary
may detect excess emission, relative to most Type Ia SNe, due to the 
collision of the SN ejecta with the non-degenerate companion 
\citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess 
emission was first detected in iPTF\,14atg  \citep{2015Natur.521..328C}, a low-velocity SN Ia with a significant and declining ultraviolet (UV) pulse detected within a
few days of the SN explosion. This UV pulse is best interpreted as a SN ejecta-companion collision (but see also \citealt{2016MNRAS.459.4428K,2017MNRAS.472.2787N}). While such emission requires a 
favorable geometric alignment and is only expected in $\lesssim$10\% of SNe Ia \citep{2010ApJ...708.1025K}, many studies have 
searched for signatures of an ejecta-companion interaction, 
typically resulting in non-detections 
(e.g., \citealt{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
  2012ApJ...744L..17B,2015Natur.521..332O,
  2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
  2015ApJS..221...22I}). Possible exceptions include SN\,2012cg, 
which exhibited excess blue emission in its early-phase light curve 
(\citealt{2016ApJ...820...92M}; though see 
\citealt{2016arXiv161007601S} for an interpretation that does not 
invoke ejecta-companion interaction), and SN\,2017cbv, which shows a clearly resolved ``bump'' in the early $UBg$ light curves \citep{2017ApJ...845L..11H}. 
% With only $\sim 10\%$ of single-degenerate progenitors occupying the preferred binary geometry for us to see the collision
% signatures, it is not surprising that there are few confirmed detections in the literature \ycao{ref?}.

Interaction is not limited to systems with a non-degenerate companion, however, as WDs enshrouded in diffuse material following a binary merger (e.g., \citealt{2015MNRAS.447.2803L}) or expanded due to a pre-explosion pulsation can give rise to ejecta-interaction signatures (e.g., \citealt{2014MNRAS.441..532D}). Models of this scenario naturally produce \ion{C}{2} absorption that is comparable in strength to \ion{Si}{2} in the days after explosion \citep{2014MNRAS.441..532D}, as was observed in SN\,2013dy \citep{2013ApJ...778L..15Z} and SN\,2017cbv \citep{2017ApJ...845L..11H}.

The vast majority of SNe Ia are observed 
to be powered purely by the radioactive decay of $^{56}$Ni. 
While the detection of SN shock cooling or ejecta interaction is rare, the level of $^{56}$Ni mixing in the SN ejecta can fundamentally alter the appearance of the SN shortly after explosion. SNe Ia experience a dark phase after
the SN shock breakout but before radioactive energy  
diffuses into the photosphere \citep{2014ApJ...784...85P}. The 
duration of this dark phase is set by how the newly synthesized 
$^{56}$Ni is mixed and deposited into different layers of the ejecta. In the case of strong mixing, the dark phase is short, or 
non-existent, as $\gamma$-rays from the radioactive $^{56}$Ni diffuse to the photosphere rapidly. This also leads to blue optical colors and a rapid initial rise in the observed light curve. If the mixing is weak and the $^{56}$Ni is confined to the innermost layers of the ejecta, the dark phase can last for several days. The early evolution of such SNe results in redder colors and a more moderate rise in luminosity \citep{2016ApJ...826...96P}. Thus, the early light curves of even non-exotic Type Ia SNe convey information about their progenitor systems by constraining the distribution of synthesized $^{56}$Ni, which in turn constrains the explosion mechanism.\footnote{\citet{2015A&A...574A..60T} found that a significant dark phase is not prevalent for a sample of Type Ib/c SNe.}

\citet{2017MNRAS.472.2787N} demonstrate that disambiguating between these different scenarios via optical photometry alone is challenging. \citeauthor{2017MNRAS.472.2787N} further show that estimates of the time of explosion, which are critical for comparing models with observations, are often incorrect by as many as $\sim$2\,d using common methods in the literature. While analytical models suggest that early spectra can be used to infer the time of explosion (e.g., \citealt{2014ApJ...784...85P}), more detailed simulations show that the photospheric evolution is not so simple \citep{2016ApJ...826...96P}. Reconciling these issues requires both a larger sample of early SNe Ia observations and more detailed models that produce synthetic light curves and spectra.

In this paper, we report observations of an extraordinarily young SN Ia, \abc, which was discovered by the intermediate Palomar Transient Factory (iPTF) on 2016 April $3.36$ UTC at $\textrm{R.A.}=13^h34^m45.49^s$, $\textrm{Dec.}=+13\degr51\arcmin14\farcs3$ (J2000) with a $g_\mathrm{PTF}$-band magnitude of $21.31\pm0.27$ \citep{2016ATel.8907....1M}. The transient is spatially coincident with a tidal tail of the galaxy NGC\,5221, which lies at a distance of $\sim$100\,Mpc. \abc\ is not detected to a limit of $g=22.1$\,mag on April $2.42$, less than 1\,d prior to discovery, and rose by $\sim$2 mag in the 24 hr following its initial detection. Our spectroscopic follow-up
campaign classified \abc\ as a normal SN Ia
\citep{2016ATel.8909....1C}. Our observations and analysis show that the early evolution of \abc\ exhibited several distinct properties relative to SN\,2011fe. We interpret those differences as arising from either strong $^{56}$Ni mixing or ejecta interaction with diffuse material, or a combination of the two.
Alongside this paper, we have released our open source analysis and all of the data utilized in this study. These are available online at \texttt{GitHub} \url{https://github.com/adamamiller/iPTF16abc}.

% This paper is organized as follows: Section \ref{sec:obs} describes
% photometric and spectroscopic observations of \abc. Section
% \ref{sec:usual_staff} establishes that \abc\ is a normal SN Ia
% in NGC\,5221. Section \ref{sec:first_light} analyzes the early
% light curve and spectra.

\section{Observations}
\label{sec:obs}

During the spring of 2016, the iPTF survey observed the field of \abc\ every night during dark time in either the $g_\mathrm{PTF}$ or 
$R_\mathrm{PTF}$ band.\footnote{P48 observations of \abc\ are reported in the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ filters throughout, which are similar to the SDSS $g'$ and Mould-$R$ filters, respectively (see \citealt{2012PASP..124..854O} for details on PTF calibration). The correction from the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ filters to SDSS $g'$ and $r'$ requires knowledge of the intrinsic source color (see Eqns.~1 and 2 in \citealt{2012PASP..124..854O}). The spectral diversity of SNe Ia in the days after explosion is poorly constrained, and as a result the color terms for \abc\ at these epochs are unknown. We proceed by assuming the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ calibration is on the AB system, which strictly speaking is incorrect, but this does not fundamentally alter any of our conclusions.} 
Survey observations were conducted with the
CFH12K camera \citep{2008SPIE.7014E..4YR} on the Palomar Observatory 
48-inch telescope (P48; \citealt{2009PASP..121.1395L}). Images were processed by the IPAC image
subtraction and discovery pipeline which subtracts off the background
galaxy light with stacked pre-SN images and performs forced
point-spread-function (PSF) photometry at the location of the SN 
\citep{2017PASP..129a4002M}. The
photometry is then calibrated to the PTF photometric catalog
\citep{2012PASP..124..854O}.

\begin{figure}[htb]
  \centering
  \includegraphics[width=0.48\textwidth]{logLC_with_colors.pdf}
  \caption{The $g$-band
  light curve of \abc. Observations from different 
  telescopes are shown with different symbols. Time is measured in 
  rest-frame days relative to the time of first light, $t_0$ (see 
  \S\ref{sec:lc_fit}). Note that the horizontal axis is shown with 
  a linear scale for $-2 \; \mathrm{d} \le t - t_0 \le 3 \; 
  \mathrm{d}$ and a $\log$ scale for $t - t_0 > 3 \; \mathrm{d}$. 
  The black ticks near the
  bottom of the panel show epochs of spectroscopic observations.
  }
  \label{fig:lightcurve}
\end{figure}

After discovery, photometric observations in the $g'$, $r'$ and $i'$
filters were obtained with the SED Machine 
(SEDm; Blagorodnova et al.\ 2017, in prep.) mounted on the Palomar Observatory 
60-inch telescope (P60). We utilized the Fremling Automated Pipeline (\texttt{FPipe}; \citealt{2016A&A...593A..68F}) to subtract galaxy light from the SEDm images using archival Sloan Digital Sky Survey (SDSS) images as a reference. This pipeline then performed forced-PSF photometry at the location of \abc, which is calibrated to the SDSS catalog \citep{2014ApJS..211...17A}.

Photometric observations in the $BVg'r'i'$ filters were conducted by the Las Cumbres Observatory (LCO) 1-m
telescope network.  PSF photometry was measured on these images using
the \texttt{lcogtsnpipe} pipeline \citep{2016MNRAS.459.3939V}. The
\textit{BV} magnitudes are calibrated to the Fourth USNO CCD
Astrograph Catalog \citep{2013AJ....145...44Z}, and the \textit{g'r'i'}
magnitudes are calibrated to SDSS Data Release 6
\citep{2008ApJS..175..297A}.

The Reionization and Transients InfraRed (RATIR) camera on the autonomous 1.5\,m Harold L. Johnson Telescope at the Observatorio Astron\'{o}mico Nacional \citep{2012SPIE.8446E..10B,2012SPIE.8444E..5LW} was used to observe \abc\ in the $r'i'ZYJH$ filters. By design, RATIR lacks a cold shutter, which means IR dark frames are not available. Laboratory testing, however, confirms that the dark current is negligible in both IR detectors \citep{2012SPIE.8453E..1OF}.

The RATIR data were reduced, coadded, and analyzed using standard CCD and IR processing techniques in \texttt{IDL}, \texttt{Python}, \texttt{SExtractor} \citep{1996A&AS..117..393B}, and \texttt{SWarp}. Aperture photometry is obtained following the methods described in \citet{2014AJ....148....2L}. The $r'i'Z$ filters are calibrated to SDSS \citep{2014ApJS..211...17A}, while the $JH$ filters are calibrated to the Two Micron All Sky Survey \citep{2006AJ....131.1163S}. For the $Y$-band calibration, we used an empirical relation in terms of the $J$ and $H$ magnitudes derived from the United Kingdom Infrared Telescope (UKIRT; \citealt{2007A&A...467..777C}) Wide Field Camera observations \citep{2009MNRAS.394..675H}.

The \textit{Swift} satellite observed \abc\ on 14 epochs, beginning 
$\sim$15\,d pre-maximum light through $\sim$22\,d post maximum. The SN 
flux is measured via aperture photometry on Ultraviolet-Optical
Telescope (UVOT) images via the usual procedures in 
\texttt{HEASoft}, including corrections for coincidence loss and 
aperture loss. The image counts
are converted to physical fluxes using the latest calibration
\citep{2011AIPC.1358..373B}. There are no pre-SN UVOT images at the 
SN location in the \textit{Swift} archive.  Visual inspection of the
UVOT images suggests negligible host-galaxy contamination in our 
UVOT flux measurements. No X-ray emission is detected from \abc\ by the \textit{Swift} X-ray Telescope (XRT).

The $g$-band discovery and follow-up data of \abc\ are 
illustrated in Figure~\ref{fig:lightcurve}.  The photometry is shown in the AB system. As previously noted the color terms necessary to convert $g_\mathrm{PTF}$ to the AB system are unknown and assumed to be zero.

Spectroscopic observations of \abc\ were taken with a variety
of telescopes and instruments over multiple epochs spanning from a
couple of days after discovery to two months after $B$-band maximum. 
An observing log is listed in Table \ref{tab:spec_obs_log}. The spectra were reduced using standard routines in \texttt{IDL}/\texttt{Python}/\texttt{Matlab}. The optical
spectral evolution of \abc\ is illustrated in Figure
\ref{fig:spec_seq}, which excludes high-resolution Very Large Telescope (VLT) spectra for clarity.

\begin{deluxetable}{cccccc}
  \tablecaption{Spectroscopic observations of \abc\ \label{tab:spec_obs_log}}
  \tabletypesize{\footnotesize}
  \tablehead{
    \colhead{Observation} & \colhead{SN} & \colhead{} &
    \colhead{} & \colhead{Range} \\
    \colhead{MJD} & \colhead{phase} & \colhead{Telescope} &
    \colhead{Instrument} & \colhead{(\AA)}
  }
  \startdata
  $57483.26$ & $-15.9$ & DCT & DeVeny\tablenotemark{1} & $3301$--$7499$ \\
  $57483.88$ & $-15.3$ & Gemini-North & GMOS\tablenotemark{2} & $3800$--$9200$ \\
  $57484.51$ & $-14.7$ & Keck-II & DEIMOS\tablenotemark{3} & $5500$--$8099$ \\
  $57486.51$ & $-12.7$ & Keck-II & DEIMOS\tablenotemark{3} & $5500$--$8099$ \\
  $57488.38$ & $-10.9$ & Keck-I & LRIS\tablenotemark{4} & $3055$--$10411$ \\
  $57489.51$ & $ -9.8$ & LCO-2m & FLOYDS\tablenotemark{5} & $3301$--$8999$ \\
  $57490.40$ & $ -8.9$ & LCO-2m & FLOYDS\tablenotemark{5} & $3301$--$9999$ \\
  $57491.55$ & $ -7.8$ & LCO-2m & FLOYDS\tablenotemark{5} & $3300$--$9998$ \\
  $57492.20$ & $ -7.2$ & VLT & X-shooter\tablenotemark{6} & $3300$--$24550$ \\
  $57494.00$ & $ -5.4$ & VLT & UVES\tablenotemark{7} & \\
  $57503.32$ & $ +3.7$ & LCO-2m & FLOYDS\tablenotemark{5} & $3300$--$9999$ \\
  $57506.00$ & $ +6.3$ & NOT & ALFOSC\tablenotemark{8} & $3602$--$8098$ \\
  $57508.27$ & $ +8.5$ & LCO-2m & FLOYDS\tablenotemark{5} & $3301$--$9999$ \\
  $57518.42$ & $+18.5$ & Keck-I & LRIS\tablenotemark{4} & $3071$--$10208$ \\
  $57520.03$ & $+20.0$ & VLT & X-shooter\tablenotemark{6} & $3300$--$24789$ \\
  $57529.40$ & $+29.2$ & LCO-2m & FLOYDS\tablenotemark{5} & $4000$--$8998$ \\
  $57542.41$ & $+41.9$ & LCO-2m & FLOYDS\tablenotemark{5} & $4000$--$8998$ \\
  $57550.40$ & $+49.7$ & LCO-2m & FLOYDS\tablenotemark{5} & $4001$--$8999$ \\
  $57562.38$ & $+61.4$ & LCO-2m & FLOYDS\tablenotemark{5} & $4800$--$9300$ \\
  \enddata
  \tablenotetext{1}{The Deveny Spectrograph \citep{2014SPIE.9147E..2NB}}
  \tablenotetext{2}{The Gemini Multi-Object Spectrograph \citep{2004PASP..116..425H}}
  \tablenotetext{3}{DEep Imaging Multi-Object Spectrograph \citep{2003SPIE.4841.1657F}}
  \tablenotetext{4}{Low-Resolution Imaging Spectrometer \citep{1995PASP..107..375O}}
  \tablenotetext{5}{FLOYDS \url{https://lco.global/observatory/instruments/floyds}}
  \tablenotetext{6}{X-shooter \citep{XShooter}}
  \tablenotetext{7}{Ultraviolet and Visual Echelle Spectrograph \citep{2000SPIE.4008..534D}}
  \tablenotetext{8}{The Andalucia Faint Object Spectrograph and Camera \url{http://www.not.iac.es/instruments/alfosc}}
\end{deluxetable}

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=0.69\textwidth]{spectra.pdf}
  \caption{Observed spectral sequence of \abc. The spectra are 
  normalized by their median flux between 6,000 and 
  7,000$\,\textrm{\AA}$.  The phase of each spectrum 
  relative to \textbf{the time of $B$-band maximum} is shown. Telluric absorption
  bands are grayed out. Line identifications are provided for 
  the spectral features discussed in the text. 
  For clarity, high-resolution spectra obtained with the VLT have been omitted (see Ferretti et al.\ 2017, submitted, for a detailed discussion of these spectra).}
  \label{fig:spec_seq}
\end{figure*}

\section{Host Galaxy, Reddening, and Classification}
\label{sec:usual_staff}

\subsection{Host Galaxy}
\label{sec:host}

The location of \abc\ is spatially coincident with a tidal tail of
galaxy NGC\,5221. \citet{2007A&A...465...71T} derived a distance
modulus of $35.0\pm0.4\,\textrm{mag}$ to NGC\,5221 from the 
Tully-Fisher relation, which is consistent with what we later derive from the SN light curve (see \S\ref{sec:classification}).

Separately, \citet{2015MNRAS.447.1531C} observe the 21-cm line in
NGC\,5221 and measure a redshift of $0.0234$, which we adopt for the remaining analysis in this paper.

\subsection{Reddening}
\label{sec:reddening}

A detailed study of the reddening towards \abc\ is presented in a 
companion paper (Ferretti et al.\ 2017, submitted). Briefly, the 
foreground Galactic extinction toward \abc\
is $E(B-V) = 0.0279 \, \mathrm{mag}$ \citep{2011ApJ...737..103S}. The high-resolution spectra presented in Ferretti et al.\ (2017) show multiple absorption components for both the \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets. While the equivalent width (EW) of these lines is quite large, implying significant extinction (e.g., \citealt{2012MNRAS.426.1465P}), Ferretti et al.\ compare the evolution of \abc\ to the well-observed normal Type Ia SN\,2011fe and only find evidence for a small amount of extinction. The empirical relation between the EW of \ion{Na}{1}\,D and extinction is known to have a large scatter, and \citet{2013ApJ...779...38P} have shown that \ion{Na}{1}\,D absorption is a poor tracer of reddening in Type Ia SNe. Thus, we adopt $E(B-V) = 0.05 \, \mathrm{mag}$ as the local extinction for \abc\ (Ferretti et al.\ 2017). For the remainder of our analysis we assume a total, Galactic $+$ host galaxy, line-of-sight extinction of $E(B-V) = 0.08 \, \mathrm{mag}$.

\subsection{Classification}
\label{sec:classification}

Using the SuperNova IDentification (\texttt{SNID};
\citealt{2007ApJ...666.1024B}) package, we find the low-resolution spectrum of
\abc\ at $+18.8$\,d is best matched by normal SNe Ia. Several characteristic
features of a SN Ia, such as \ion{Si}{2}, \ion{S}{2}, can be easily identified
in the spectra of \abc\ (Figure \ref{fig:spec_seq}). From the $+3.7 \,
\mathrm{d}$ LCO spectrum, we measure the pseudo-equivalent widths (pEWs) of
the absorption features near 5750 and 6100 \AA, which are respectively
attributed to \ion{Si}{2}\,$\lambda\lambda$6580, 7234, to be $-12 \pm 2$\,\AA\
and $-55 \pm 5$\,\AA, respectively. According to the classification scheme
presented in \citet{2006PASP..118..560B}, these pEW measurements indicate that
\abc\ is a shallow silicon SN, similar to the SN\,1999aa-likes
\citep{2009PASP..121..238B}. \textbf{In Figure~\ref{fig:branch_vel}, the
velocity evolution of \ion{Si}{2}$\,\lambda6355$ in \abc\ is compared to the
median evolution of the four spectroscopic subclasses defined in
\citet{2006PASP..118..560B}. The median evolution is defined using the sample
of SNe Ia in \citet{2012AJ....143..126B}. For each SN, we interpolate the
\ion{Si}{2} velocity, $v_{\mathrm{Si\,II}\,\lambda6355}$, to a fixed grid at
1\,d intervals. We then take the median $v_{\mathrm{Si\,II}\,\lambda6355}$ for
each point on the grid, and only retain those points with at least 3 SNe (to
prevent the early and late evolution to be defined by a single SN). The
evolution of \ion{Si}{2} is most reminicent of the shallow-silicon subclass
defined in \citet{2006PASP..118..560B}. }

\begin{figure}[htb]
  \centering
  \includegraphics[width=0.48\textwidth]{Branch_vel_evol.pdf}
  \caption{\textbf{Velocity evolution of \ion{Si}{2},   
           $v_{\mathrm{Is\,II}\,\lambda6355}$, for \abc\ compared to the median
           evolution of the four spectroscopic subclasses defined in
           \citet{2006PASP..118..560B}, using data from
           \citet{2012AJ....143..126B}. Typical uncertainties for \abc\ are
           $\sim$1000\,km\,s$^{-1}$ before $T_{B,\mathrm{max}}$, when the
           \ion{Si}{2}$\,\lambda6355$ profile is shallow and the minimum of
           absorption is difficult to determine, and $\sim$300\,km\,s$^{-1}$
           after $T_{B,\mathrm{max}}$. For the median curves, the typical
           scatter, determined via the interquartile range of the sample, is
           $\sim$700\,km\,s$^{-1}$, around $T_{B,\mathrm{max}}$.
           At early times, core-normal and broad-line SNe have significantly
           faster \ion{Si}{2} than \abc, while the declining trend of coolSNe
           does not match \abc.}
  }
  \label{fig:branch_vel}
\end{figure}

To determine the brightness and time of $B$-band maximum for 
\abc, we fit the P60 light curves with the \texttt{sncosmo} software package.\footnote{\texttt{sncosmo} is available at
  \url{https://sncosmo.readthedocs.io}.} This fit includes a \texttt{SALT2} template \citep{2007A&A...466...11G} that has been corrected for extinction using the \citet{1999PASP..111...63F} reddening law, $R_V=3.1$, and $E(B-V) = 0.08 \; \mathrm{mag}$.

We determine the time of rest-frame \textit{B}-band maximum to be
$\textrm{MJD}_\mathrm{max}=57499.54\pm0.23$, the coefficient of the zeroth
principle component $x_0 = 0.0086 \pm 0.0003$, the coefficient of the first
principle component $x_1 = 0.96 \pm 0.15$, and the color term $c = 0.033 \pm
0.029$. The best-fit model also gives an unreddened apparent peak magnitude of
$m^*_{B}=15.80 \pm 0.04 \,\textrm{mag}$ in the SN rest frame. In the following
sections, we define the best-fit value $\textrm{MJD}_\mathrm{max}=57499.54$ as
the time of $B$-band maximum, $T_{B_\mathrm{max}}$, which we also define as
phase $t=0$.

We measure the (pseudo)bolometric luminosity, $L_{UVOIR}$, of \abc\ at peak
via trapezoidal integration of the reddening-corrected flux from the UV,
optical, and NIR (UVOIR) filters. The light curves in the individual filters
are interpolated so that $L_{UVOIR}$ is evaluated at common epochs in each
filter. From this integration, we determine that \abc\ reaches a maximum
luminosity $L_\mathrm{max} = 1.2 \pm 0.1 \times 10^{43} \; \mathrm{erg \;
s}^{-1}$. This value is consistent with the normal SNe Ia studied in
\citet{2016A&A...588A..84D}, a sample which excludes low-luminosity SNe.
\textbf{Following Arnett's rule \citet{1982ApJ...253..785A,
1985Natur.314..337A}, it is possible to derive the mass of $^{56}$Ni
synthesized in the explosion from $L_\mathrm{max}$. Assuming a rise time of
$19 \pm 3$\,d (see \citealt{2006A&A...450..241S}) we find $M_\mathrm{Ni} = 0.6
\pm 0.1 M_\odot$.\footnote{Assuming a rise time of 17.9\,d, see
\S\ref{sec:lc_fit} results in a consistent estimate of $M_\mathrm{Ni}$.}}

After establishing \abc\ as a normal SN Ia, we use the latest calibration
\citep{2014A&A...568A..22B} of the Phillips relation
\citep{1993ApJ...413L.105P} using $m^*_{B}$, $x_1$ and $c$ to derive a
distance modulus $\mu = 34.89 \pm 0.10 \,\textrm{mag}$ to the SN, provided
that the host galaxy of \abc\ has a stellar mass $< 10^{10}\sm$. A more
massive host galaxy would result in a larger inferred distance modulus that is
nevertheless consistent within the uncertainties. For the following analysis
we adopt a distance modulus $\mu = 34.89 \pm 0.10 \,\textrm{mag}$ for
\abc.\footnote{This $\mu$ is consistent with the $z_\mathrm{SN}$, $H_0 = 73 \,
\mathrm{km \, s^{-1} \, Mpc}^{-1}$, and Virgo-infall-corrected distance
\citep{2000ApJ...529..786M}.}

\section{Early Observations}
\label{sec:first_light}

Here we consider our suite of early observations of \abc, and compare our
findings with SN\,2011fe, a well-studied, nearby SN that was discovered
shortly after explosion
\citep{2011Natur.480..344N,2012ApJ...744L..17B,2014ApJ...784...85P}.

\subsection{Time of First Light from the Early Light Curve}
\label{sec:lc_fit}

The time of first light for SNe is usually estimated by extrapolating
early-phase light curves to determine when the SN flux is equal to 0.
Assuming an ideal, expanding fireball with constant temperature,
\citet{1982ApJ...253..785A} derives that $f \propto t^2$, where $f$ is the SN
flux and $t$ is the time since explosion. Despite the basic assumption of a
constant temperature at early times, multiple studies have found that the
early emission from Type Ia SNe can be described as a power law in time, with
power-law index consistent with 2, i.e. $f \propto t^2$ (e.g.,
\citealt{2006AJ....132.1707C, 2010ApJ...712..350H,
2011MNRAS.416.2607G}).\footnote{Many of the studies that find $f \propto t^2$
sample SNe Ia light curves at a phase that is closer to $T_{B,\mathrm{max}}$
than the initial observations of \abc\ discussed here.}

As our observations include especially early observations of \abc, there are
upper limits $\sim$1 d prior to the discovery epoch, we model the early flux
from \abc\ as a power law, but allow the power-law index to vary:

\begin{equation}
  \label{eq:broken_power_law}
  f(t) \left\{
    \begin{array}{ll}
      = 0,\ \textrm{when}\ t \le t_0 \\
      \propto (t-t_0)^{\alpha},\ \textrm{when}\ t>t_0
    \end{array}
  \right.\ ,
\end{equation}
where $t_0$ is the time of first light, $\alpha$ is the power-law index, and
$t$ is measured in the SN rest-frame. \textbf{We allow $\alpha$ to vary to
find the best match to the data, and later show that models with $\alpha = 2$
do not explain the data.} To determine $t_0$ and $\alpha$ we fit the earliest
observations of \abc. Due to slight variations in the passbands, we fit the
model only to the relative flux measured in the $g_\mathrm{PTF}$-band, which
is the only filter with observations prior to first light, a necessity for
constraining $t_0$.

To determine the best fit parameters, we search a large grid over $t_0$,
$\alpha$, and the proportionality constant, and minimize $\chi^2$. The
modeling results show that the SN flux rises approximately linearly between
$t=-18\,\textrm{d}$ and $t=-15\,\textrm{d}$. Figure \ref{fig:early_lc_fit}
shows the best-fit result and the joint marginal distribution of $t_0$ and
$\alpha$. From the best-fit model we obtain $\alpha=0.98 \pm ^{0.16}_{0.14}$
and $t_0=-17.91 \pm ^{0.07}_{0.15}$ d, where the uncertainties represent the
marginalized 95\% confidence intervals. Our first detection of \abc\ occurred
$\sim$${0.15}\,\textrm{d}$ after the SN first light. In the analysis that
follows, the precise values of the best-fit parameters are not important. The
critical finding here is that $\alpha \approx 1$ and $t_0 \approx -18 \,
\mathrm{d}$.

Figure \ref{fig:early_lc_fit} also shows the best-fit model while fixing
$\alpha = 2$. The $f \propto t^2$ model does not match the observations.
Formally, for the $\alpha = 2$ model $\chi^2 = 63.7$ with $\nu = 15$ degrees
of freedom (dof), while $\chi^2 = 10.2$ with $\nu = 14$ dof for the $\alpha =
0.98$ model.

As previously noted, a precise determination of the rise time,
$t_\mathrm{rise}$, of SNe Ia is challenging as there may be a dark phase
following explosion \citep{2014ApJ...784...85P}. Nevertheless, to be
consistent with previous studies (e.g., \citealt{2011MNRAS.416.2607G}), here
we find $t_\mathrm{rise} = 17.91\pm _{0.07}^{0.15}$ d based on our fit for
$t_0$. We caution that this measurement may be underestimated due to the
uncertain time of explosion.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.48\textwidth]{early_lc.pdf}
  \caption{Best-fit power-law model to describe the early flux 
  from \abc\ in the $g_\mathrm{PTF}$ band. \textit{Top}: The 
  relative flux, 
  $f_{g_\mathrm{PTF}}$, shown as green circles, is measured via 
  forced-PSF photometry at the SN location. The model flux, 
  adopting best-fit parameters $\alpha=0.98$ and 
  $t_0=-17.94\,\textrm{d}$, is shown as a dashed black line. Also 
  shown is the best-fit model after fixing 
  $\alpha=2$ (thin, dash-dot line). 
  The joint distribution of $t_0$ and $\alpha$ for the best-fit 
  power-law model is illustrated in the inset. 
  The solid contours represent the $68\%$ and $99.7\%$ confidence 
  levels.
  \textit{Bottom}: 
  The observations and models following 
  subtraction of the best-fit power-law model, $f_{\alpha = 0.98}$.
  The $t^{0.98}$ model provides a much better fit to the 
  observations than the $t^2$ model. 
  }
  \label{fig:early_lc_fit}
\end{figure}

Simulations presented in \citet{2017MNRAS.472.2787N} show that the flux from
SNe Ia explosion models does not evolve as a power law in time. Despite this
non-power-law evolution, \citet{2017MNRAS.472.2787N} show that $f \propto
t^\alpha$ models produce good formal fits to simulated observations. The $f
\propto t^\alpha$ fits to the \citeauthor{2017MNRAS.472.2787N}\ models result
in glaring errors to the estimated explosion times. Despite the caution
advised in \citet{2017MNRAS.472.2787N}, we proceed with the power-law fits
from above for several reasons. (1) Our primary aim with the fit is to
characterize $\alpha$ for \abc\ compared to SN\,2011fe (see below). (2) The
errors in the estimated explosion times from \citeauthor{2017MNRAS.472.2787N}
increase the later the first observation of the SN is obtained.
\citeauthor{2017MNRAS.472.2787N} only present fits to observations that start
$>$2\,d after explosion, whereas the discovery epoch for \abc\ is likely much
earlier than that (see \S\ref{sec:early_vel}). (3) We do not use the fits to
estimate the time of explosion, $t_0$ instead is an estimate of the time of
first light. \textbf{Finally,} (4) Pre-discovery upper limits, which are not
included in \citeauthor{2017MNRAS.472.2787N}, provide additional strong
constraints on the time of first light.

Unlike \abc, the early emission from SN\,2011fe is well fit by a $f \propto
t^2$ model \citep{2011Natur.480..344N}. Thus, the near-linear rise in flux
over the first few days after first light for \abc\ is distinct compared to
SN\,2011fe. To our knowledge this behavior has only been observed in 2 other
SNe (2013dy, 2014J;
\citealt{2013ApJ...778L..15Z,2014ApJ...783L..24Z,2015ApJ...799..106G}). Any
model to explain the observations of \abc\ must account for this near-linear
rise in the days after first light.

\subsection{Time of Explosion from the Photospheric Velocity}
\label{sec:early_vel}

The potential dark phase following shock breakout, where SNe Ia are not luminous because the energy deposition from radioactive decay has not had sufficient time to diffuse to the photosphere, means that the explosion time $t_\mathrm{exp} \ne t_0$ in some cases. \citet{2014ApJ...784...85P} suggest that measurements of the photospheric velocity be used to determine $t_\mathrm{exp}$ given that the ejecta begin expanding from the moment of explosion.  Assuming a constant opacity in the ejecta, \citeauthor{2014ApJ...784...85P} 
find that the photospheric velocity evolves as
$v_\mathrm{ph}\propto(t-t_\mathrm{exp})^{-0.22}$. Numerical experiments by \citet{2016ApJ...826...96P} find that the constant-opacity assumption strongly depends on the amount of $^{56}$Ni mixing in the SN ejecta (for a more detailed discussion of $^{56}$Ni mixing see \S\ref{sec:Ni_mixing}). As a result, the adoption of a $t^{-0.22}$ power-law model may not be valid for all SNe Ia. Nevertheless, we proceed on the assumption that \abc\ experienced strong $^{56}$Ni mixing, corresponding to the models that are best approximated as a $t^{-0.22}$ power law. We do this in part to compare with previous studies, though we caution that the inferred value of $t_{exp}$ is subject to uncertainties related to ejecta mixing.

While the photospheric velocity
is not easy to measure, line velocities of \ion{Si}{2} or \ion{Ca}{2}
can be used as a proxy
\citep{2014ApJ...784...85P,2016ApJ...826..144S}. In the case of \abc, the \ion{Ca}{2} IR triplet is very weak, 
likely due to high temperatures in the ejecta. Thus, we
determine the photospheric velocity from the 
\ion{Si}{2}\,$\lambda$6355 line. Visual inspection shows no 
sign of multi-velocity components of \ion{Si}{2}, and that 
the \ion{C}{2}\,$\lambda$6580 line overlaps the red
wing of the \ion{Si}{2} line (see Figures~\ref{fig:spec_seq} and \ref{fig:carbon}). Consequently, we model the observed 
spectra between 5900 and 6500\,\AA\ (rest-frame) as the 
combination of two Gaussian kernels plus a linear baseline, which
accounts for \ion{Si}{2}, \ion{C}{2} and the
continuum, respectively. The expansion
velocity of \ion{Si}{2} is measured by the central wavelength
of the \ion{Si}{2} Gaussian kernel.

\begin{figure*}[!thb]
  \centering
  \includegraphics[width=0.48\textwidth]{Chi2.pdf}
  \includegraphics[width=0.48\textwidth]{VelocityPlot.pdf}
  \caption{Constraints on $t_\mathrm{exp}$ from fitting the velocity
    evolution of \ion{Si}{2}.
    \textit{Left panel:} the dashed, solid and dash-dotted curves 
    show $\chi^2$ for fitting power laws with
    indices $-0.20$, $-0.22$ and $-0.24$, respectively. The blue
    vertical line and the orange shaded region indicate $t_0$ and 
    its 95\% confidence interval from Section
    \ref{sec:lc_fit}, respectively.
    \textit{Right panel:} Observed \ion{Si}{2}\,$\lambda$6355
    velocities (blue circles) and the best-fit power-law model
     with an index of $-0.22$ (dashed line). Additionally, the 
     measured velocities of \ion{C}{2}\,$\lambda\lambda$6580, 7234 
     are shown. Typical uncertainties are smaller than the size of 
     the points. \textit{Right inset}: Evolution of the 
     pseudo-equivalent width of \ion{Si}{2}\,$\lambda$6355, and 
     \ion{C}{2}\,$\lambda\lambda$6580, 7234 in the $\sim$7~d    
     following explosion.}
  \label{fig:velocity_t_exp}
\end{figure*}

We fit the measured velocities of \ion{Si}{2}\,$\lambda$6355 to the
$v\propto(t-t_\mathrm{exp})^{-0.22}$ model by minimizing the $\chi^2$ value and find the best-fit explosion time relative to $T_{B,\mathrm{max}}$ in the SN rest frame to be $t_\mathrm{exp} = -17.45 \pm ^{0.14}_{0.16}\,\textrm{d}$, where the uncertainties represent the 95\% confidence interval (Figure \ref{fig:velocity_t_exp}). Following the analysis in \citet{2014ApJ...784...85P}, we additionally alter the power-law index to $-0.20$ and $-0.24$ to examine the sensitivity of the result on the assumed power-law index. We find that this variation in the power-law index results in a change of $t_\mathrm{exp}$ of $\approx\pm$0.5\,d (Figure \ref{fig:velocity_t_exp}). Given the analytical approximation that $v \propto t^{-0.22}$, we adopt $t_\mathrm{exp} = -17.5 \pm 0.5 \; \mathrm{d}$, where the uncertainty reflects possible variations in the power-lax index (see \citealt{2014ApJ...784...85P}).

Comparing our estimates for $t_\mathrm{exp}$ and $t_0$ (left panel of 
Figure \ref{fig:velocity_t_exp}), we find that 
$t_0\lesssim t_\mathrm{exp}$.
Since physical causality requires $t_\mathrm{exp} \le t_0$, we draw the qualitative
conclusion that $t_0\simeq t_\mathrm{exp}$, which is consistent to within the uncertainties. Again, we caution that this derivation of $t_\mathrm{exp}$ relies on the assumption $v \propto t^{-0.22}$, which may not be valid for all SNe Ia.

% From $t_{exp}$, we estimate the actual rise time of \abc\ from the
% time of explosion to \textit{B}-band maximum to be $17.95$ d.  In
% comparison, SN2011fe has $t_0=-17.7$\,d \citep{2013A&A...554A..27P}
% with a preceding dark phase of $\sim 1$ day
% \citep{2014ApJ...784...85P}. ASASSN-14lp has $t_0=-16.94$ days with a
% preceding dark phase of about acouple of days
% \citep{2016ApJ...826..144S}. This comparison suggests that the actual
% rise time of a SN Ia between the SN explosion and the \textit{B}-band
% maximum is roughly a constant, which largely depends on the total mass
% of synthesized $^{56}$Ni.\ycao{I am not convinced by this last sentence}

\subsection{Strong and Short-Lived Carbon Features}
\label{sec:carbon}

The early spectra of \abc\ exhibit unusually strong absorption due 
to \ion{C}{2} $\lambda\lambda$6580, 7234. We highlight the evolution 
of these spectral features in Figure~\ref{fig:carbon}. From these 
spectra we see that \ion{C}{2}\,$\lambda$6580 is as strong as 
\ion{Si}{2}\,$\lambda$6355 at $t \approx -15 \; \mathrm{d}$. The 
strength of the \ion{C}{2} lines declines with time, and by $t \approx -10 \; \mathrm{d}$ \ion{C}{2} is no longer detectable.

\begin{figure}[]
  \centering
  \includegraphics[width=0.49\textwidth]{CarbonFeature.pdf}
  \caption{Evolution of the \ion{C}{2} features observed in the 
  early spectra of \abc. The raw spectra are shown in orange, while 
  the solid blue lines show the best-fit models including Gaussian 
  components for each line and a linear component for the continuum 
  (see text for further details). The dark grey vertical lines show 
  the measured line centers, indicating the decline in the 
  photosphere velocity in the $\sim$7~d after explosion (note that 
  \ion{C}{2}\,$\lambda$7234 is not detected in the $-10.9$\,d 
  spectrum). The phase of each spectrum relative to 
  $T_{B,\mathrm{max}}$ is labeled.
  }
  \label{fig:carbon}
\end{figure}

Similar to our analysis of the \ion{Si}{2}\,$\lambda$6355 line, we 
can measure velocities and pEWs of 
\ion{C}{2}\,$\lambda\lambda$6580, 7234. We compare the velocity 
evolution of the \ion{C}{2} lines to \ion{Si}{2} in the right panel 
of Figure \ref{fig:velocity_t_exp}, which also shows the pEW evolution of these lines. These measurements confirm the qualitative analysis from Figure~\ref{fig:carbon}: namely, the pEW of \ion{C}{2}\,$\lambda$6580 is comparable to that
of \ion{Si}{2}\,$\lambda$6355 at $t \approx -16 \; \mathrm{d}$ and 
the pEW of \ion{C}{2} 
decreases until the feature is no longer detectable 
around $t = -10$ d.

The detection of \ion{C}{2} in SN Ia spectra is relatively rare
as it requires both unburned carbon, which is likely only present in the outermost layers of the ejecta, and non-local thermal equilibrium effects in order to excite the ionized carbon (e.g., \citealt{2007ApJ...654L..53T}). Spectra obtained around or after $T_{B_\mathrm{max}}$ rarely show \ion{C}{2} as the photosphere has receded from the outermost ejecta, while pre-max spectra show evidence for \ion{C}{2} in $\sim$1/4 of all normal SNe Ia (e.g., \citealt{2011ApJ...732...30P,2012MNRAS.425.1917S,2011ApJ...743...27T}), but the signatures are typically weak. While we caution that the sample of SNe Ia with spectra taken within a few days of explosion is small, SNe\,2013dy and 2017cbv are the only other objects known to have strong \ion{C}{2} features like \abc\ \citep{2013ApJ...778L..15Z,2017ApJ...845L..11H}. As a counterexample, SN\,2011fe only exhibited weak \ion{C}{2} features in its first spectra
\citep{2012ApJ...752L..26P}. Thus, models of \abc\ must explain the strong \ion{C}{2} absorption observed shortly after explosion.

\subsection{Blue Optical Colors Shortly After Explosion}

Multiband observations of \abc\ began $\sim$1.5 d after discovery, which allows us to trace its color evolution starting $\sim$1.7 d after $t_0$. In Figure~\ref{fig:B-Vcolors} we compare the $(B - V)_0$ color evolution of \abc\ and SN\,2011fe, where the observations of SN\,2011fe are taken from \citet{2016ApJ...820...67Z}. For both SNe the colors have been corrected for the total inferred reddening along the line of sight. Interestingly, \abc\ has a nearly flat color evolution up to 10\,d before $T_{B,\mathrm{max}}$, while SN\,2011fe exhibits red colors initially before evolving to the blue.

\begin{figure}[]
  \centering
  \includegraphics[width=0.48\textwidth]{16abc_11fe_colors.pdf}
  \caption{$(B - V)_0$ color evolution of \abc\ (shown as squares) 
    compared to SN\,2011fe (shown as stars). The $B$ and $V$ 
    photometry are calibrated on the Vega system, and have been 
    corrected for extinction. The data for 
    SN\,2011fe are from \citet{2016ApJ...820...67Z}.}
  \label{fig:B-Vcolors}
\end{figure}

Roughly 16\,d prior to $B$-band maximum, the $(B - V)_0$ color of \abc\ is $\sim$0.5 mag bluer than SN\,2011fe. Like \abc, SN\,2012cg \citep{2016ApJ...820...92M} and SN\,2017cbv \citep{2017ApJ...845L..11H} exhibit $(B - V)_0$ colors that are significantly bluer than SN\,2011fe at very early epochs. While there are many factors that contribute to the early optical colors of SNe Ia (see \S\ref{sec:lc_energy} below), early blue colors are often interpreted as a hallmark of interaction between the SN ejecta and a binary companion. Nevertheless, it is interesting to note that despite the blue optical colors, the $\mathrm{UV} - \mathrm{optical}$ colors of \abc, SN\,2012cg, and SN\,2017cbv are significantly redder at these early epochs than the $\mathrm{UV} - \mathrm{optical}$ colors of iPTF\,14atg \citep{2015Natur.521..328C}, the most likely candidate for SN ejecta-companion interaction.

\section{Discussion: \textbf{Modeling the Early Evolution of \abc}}
\label{sec:lc_energy}

Relative to the nearby, normal SN\,2011fe, we have identified several distinct characteristics of the early evolution of \abc, including: (i) a near-linear photometric rise, (ii) a qualitatively short, or possibly absent, dark phase, assuming $v_\mathrm{ph} \propto t^{-0.22}$, (iii) the presence of strong \ion{C}{2} absorption, and (iv) blue and non-evolving $(B - V)_0$ color in the week after explosion. While most SNe Ia are powered purely by radioactive decay, the observed radiation shortly after explosion can also include contributions 
from SN shock cooling or the collision of the SN ejecta with a non-degenerate companion or \textbf{nearby, unbound} material. Here we consider those scenarios as a possible explanation for the early behavior of \abc.

\subsection{SN Shock Cooling}

The shock breakout of a SN Ia lasts for a fraction of a second due to
compact size of the exploding star. Emission from the subsequent 
cooling phase may last for several days, however (e.g., 
\citealt{2010ApJ...708..598P}). Following the analysis of \citet{2012ApJ...744L..17B} for SN\,2011fe, we
compare the early-phase $g_\mathrm{PTF}$ light curve of \abc\ with
two shock cooling models \citep{2011ApJ...728...63R, 
2010ApJ...708..598P}. From this analysis, we constrain the \abc\ 
progenitor radius to be $<1\,\sr$. Our observations of \abc\ cannot 
place tight constraints on the size of SN Ia progenitors. Indeed, 
for a typical WD radius, such as that inferred for SN\,2011fe 
($\lesssim 0.02$--$0.04\,\sr$; \citealt{2012ApJ...744L..17B, 
2014ApJ...784...85P}), the expected emission from shock cooling is 
$\sim$2 mag fainter than the P48 $g_\mathrm{PTF}$ detection limit at this distance. 
Thus, we conclude that shock cooling does not contribute to the 
early emission from \abc.

\subsection{SN-Companion collision}
\label{sec:companion}

The detection of emission from the collision of the SN ejecta 
with a non-degenerate companion requires a favorable orbital 
alignment relative to the line of sight. Thus, from geometric 
considerations alone the probability of detecting ejecta-companion 
interaction is low, $\sim$10\%. \citet{2010ApJ...708.1025K} 
calculates that the collision of SN ejecta with a companion  
generates thermal emission with a spectrum that peaks in the 
UV. The resulting \textit{g}-band emission is expected to be weak.

To examine the possibility of a SN-companion signature in the early
light curve of \abc, we employ the \citet{2010ApJ...708.1025K}
model and assume canonical values for the ejecta mass, $1.4\,\sm$, expansion velocity, $10^{4}\,\textrm{km}\,\textrm{s}^{-1}$, and a constant opacity, $0.2\,\textrm{cm}^2\,\textrm{g}^{-1}$. We calculate the expected $g_\mathrm{PTF}$ brightness of an ejecta-companion collision at the distance of \abc\ behind a total reddening of $E(B-V) = 0.08 \, \mathrm{mag}$ using the parameterized equations in \citet{2012ApJ...749...18B}. If we assume the binary is aligned with the optimal orientation relative to the line of sight, a binary separation of $a \approx 3 \times 10^{11}\, \mathrm{cm}$ is needed to explain the initial detection of \abc, as shown in Figure~\ref{fig:SN-companion}.  The minimum binary separation capable of explaining the observed brightness at the epoch of discovery is $a \approx 10^{11} \, \mathrm{cm}$. Figure~\ref{fig:SN-companion} shows that such models peak at $g_\mathrm{PTF} \approx 21.5 \, \mathrm{mag}$, provided that $t_\mathrm{exp} \approx t_0 - 0.3 \, \mathrm{d}$. These models do not, however, match the $g_\mathrm{PTF}$ evolution for $t > t_0 + 0.5 \, \mathrm{d}$ (though it is possible that the SN photosphere dominates the companion-interaction signature at this phase). While they are otherwise compatible with the observations, we do not favor the above models as the explanation for the early flux from \abc\ because they do not explain the 2 mag rise in the $\sim$24 hr after discovery.

\begin{figure}[!thb]
  \centering
  \includegraphics[width=0.48\textwidth]{companion_lc.pdf}
  \caption{Comparison of SN ejecta-companion interaction models 
  with early observations of \abc. $g_\mathrm{PTF}$ 
  detections and 3$\sigma$ upper limits are shown 
  as green circles and downward arrows, respectively. 
  \textit{Swift}/$UVW1$ observations are shown as magenta hexagons.
  The dashed, solid, and dash-dot lines show the expected flux 
  for companion-interaction models in the $g_\mathrm{PTF}$ (green) 
  and $UVW1$ (magenta) filters. The models have been adjusted to 
  account for the distance and reddening towards \abc. Each 
  model features a different companion semi-major axis, $a$, and 
  time of explosion, $t_\mathrm{exp}$, as labeled in the 
  legend. While models with $a \gtrsim 10^{12} \, \mathrm{cm}$ can 
  explain the early optical rise, they greatly 
  overpredict the UV flux. Models with $a \approx 10^{11} \, 
  \mathrm{cm}$ can explain the initial detection \abc, but they 
  fail to replicate the $\sim$2\,mag rise in the $\sim$24\,hr 
  after explosion.
  }
  \label{fig:SN-companion}
\end{figure}

Figure~\ref{fig:SN-companion} additionally shows that a companion at $a \approx 18 \times 10^{11} \, \mathrm{cm}$ provides a good match to the initial optical rise, if $t_\mathrm{exp} \approx t_0 + 0.1 \; \mathrm{d}$. Models with $a \gtrsim 10^{12} \, \mathrm{cm}$, which can explain the initial $g_\mathrm{PTF}$ rise, significantly overpredict the observed UV flux, however. There is no choice of $a$ capable of replicating the early rise of \abc\ without also overpredicting the observed UV flux.

The challenges associated with each of the previously considered models lead us to conclude that the early evolution of \abc\ cannot be explained via the interaction of the SN ejecta with a companion. We cannot, however, exclude the presence of a red-giant, or other non-degenerate, companion as our calculations have assumed that the binary is aligned with the optimal geometry relative to the line of sight. If the geometry is not favorable, then it is possible that signatures from interaction with a companion are not visible.

\subsection{\textbf{Sub-Chandrasekhar Detonations and Pure Deflagrations}}

\textbf{In \citet{2017MNRAS.472.2787N} the early photometric evolution of Type Ia SNe is explored via a variety of explosion models and detailed radiative-transfer calculations. Specifically, \citet{2017MNRAS.472.2787N} examine 2 Chandrasekhar-mass ($M_\mathrm{Ch}$) explosions and compare their evolution to sub-Chandrasekhar explosions and pure deflagrations. The 2 $M_\mathrm{Ch}$ explosions include the ``W7'' carbon-deflagration model of \citet{1984ApJ...286..644N} and the ``N100'' delayed-detonation model from \citet{2013MNRAS.429.1156S}. The sub-Chandrasekhar models include: a violent WD-WD merger, which triggers a carbon detonation in the more massive WD, a centrally ignited detonation, and a sub-Chandrasehkar double detonation explosion, in which a He-surface-layer detonation triggers a carbon detonation in the WD core. \citet{2017MNRAS.472.2787N} note that, of these last two sub-Chandrasehkar models, the later provides the more realistic scenario. Finally, \citet{2017MNRAS.472.2787N} also examine the ``N5def'' and ``N1600Cdef'' pure deflagrations from \citet{2014MNRAS.438.1762F}.}

\begin{figure}[]
  \centering
  \includegraphics[width=0.48\textwidth]{compare_Noebauer.pdf}
  \caption{
  \textbf{Comparison of the models from \citet{2017MNRAS.472.2787N}
  to the \abc\ $g$-band light curve (\textit{top}) and the $(B-V)_0$ color curve (\textit{bottom}). The
  light curve has been corrected for the distance modulus to \abc\
  (\S\ref{sec:host}), while both the light curve and color curve have been
  corrected for reddening. Each light curve is translated to match the LCO
  observations $\sim$1.7\,d after $t_0$ given the uncertain time of explosion.
  Translational offsets are listed in the bottom panel legend, which shows the
  models in order of decreasing $M_\mathrm{Ni}$ from top to bottom. DD models
  (labeled as DDC to match the nomenclature of \citealt{2014MNRAS.441..532D})
  are shown as dash-dot lines, while PDD models (labeled as PDDEL) are shown
  as dashed lines. The inset in the top panel shows the residuals relative to
  the PDDEL7n model. The PDD models provide a better match to the
  observations.} 
  } \label{fig:noebauer} 
\end{figure}

\textbf{In Figure~\ref{fig:noebauer} we compare photometric observations of
\abc\ to the large-$^{56}$Ni yield DD and PDD models presented in
\citet{2014MNRAS.441..532D}.
}

The detonation of a WD with $M < 1.4 \, M_\odot$, which can occur if triggered by a He-surface-layer detonation (known as the sub-Chandrasekhar double detonation), can produce a rapid rise and early blue optical colors (e.g., \citealt{2017MNRAS.472.2787N}). While this qualitative behavior matches \abc, the $U-V$ colors predicted in \citet{2017MNRAS.472.2787N} do not match our observations. In particular, the sub-Chandrasekhar double detonation model $U - V$ colors rapidly evolve from $\sim$$-0.6\,\mathrm{mag}$ $\sim$1\,d after explosion to $\sim$0.5\,mag $\sim$3\,d after explosion. We observe a steady $U - V$ color evolution from the red to the blue for \abc, however, with $U - V \approx 0.6 \, \mathrm{mag}$ $\sim$2\,d after $t_0$ and $U - V \approx 0 \, \mathrm{mag}$ $\sim$8\,d after $t_0$. As such we conclude that \abc\ is not the result of a sub-Chandrasekhar double detonation. \citet{2017MNRAS.472.2787N} note that their one-dimensional models may not capture the predicted asphericity for sub-Chandrasekhar double detonations, though moving away from the preferred viewing angle results in redder colors and a more gradual rise, which is not observed for \abc.

\subsection{Interaction with \textbf{Nearby, Unbound} Material}

To model SN\,2011fe, \citet{2014MNRAS.441..532D} recently examined
pulsational delayed-detonation (PDD) models as an explanation for some SNe
Ia. Briefly, PDD models differ from ``standard'' delayed detonation (DD)
models in that the expansion of the WD during the initial deflagration
phase leads to the release of unbound material. Following this pulsation,
the bound material contracts eventually triggering a subsequent
detonation.\footnote{\citet{2014MNRAS.441..532D} note that the deflagration
and detonation in their PDD models are artificially triggered.} An
important consequence of this progression for PDD models is that the
unbound material expands and avoids burning, unlike DD models that
typically leave no unburnt material. This results in significantly more
carbon in the outer layers of the SN ejecta \citep{2014MNRAS.441..532D}.

\textbf{\citet{2014MNRAS.441..532D} find that DD models are universally faint and red at early times, $\sim$24--48\,hr after explosion, while the PDD models exhibit a faster rise and bluer colors. Briefly, this occurs in the PDD scenario because the collision with the unbound material surrounding the WD heats the outer layers of the SN ejecta. Importantly, the PDD models are nearly indistinguishable from DD models around and post peak.}

\begin{figure}[]
  \centering
  \includegraphics[width=0.48\textwidth]{compare_Dessart.pdf}
  \caption{
  \textbf{Comparison of the DD and PDD models of \citet{2014MNRAS.441..532D}
  to the \abc\ $g$-band light curve (\textit{top}) and the $(B-V)_0$ color curve (\textit{bottom}). The
  light curve has been corrected for the distance modulus to \abc\
  (\S\ref{sec:host}), while both the light curve and color curve have been
  corrected for reddening. Each light curve is translated to match the LCO
  observations $\sim$1.7\,d after $t_0$ given the uncertain time of explosion.
  Translational offsets are listed in the bottom panel legend, which shows the
  models in order of decreasing $M_\mathrm{Ni}$ from top to bottom. DD models
  (labeled as DDC to match the nomenclature of \citealt{2014MNRAS.441..532D})
  are shown as dash-dot lines, while PDD models (labeled as PDDEL) are shown
  as dashed lines. The inset in the top panel shows the residuals relative to
  the PDDEL7n model. The PDD models provide a better match to the
  observations.} 
  } \label{fig:dessart} 
\end{figure}

\textbf{In Figure~\ref{fig:dessart} we compare photometric observations of
\abc\ to the large-$^{56}$Ni yield DD and PDD models presented in
\citet{2014MNRAS.441..532D}. DD and PDD models with $M_\mathrm{Ni}/M_\odot \la
0.7$ and 0.5, respectively, fail to match the early luminosity and blue colors
of \abc. The PDD models provide a better match to the observations than the DD
models. The PDD models evolve more rapidly towards blue colors and provide a
better match to the $g$-band flux in the days after explosion. Furthermore,
unlike the DD models, the PDD models exhibit strong \ion{C}{2} lines that
gradually disappear in the $\sim$week after explosion
\citet{2014MNRAS.441..532D}. In detail, the early wiggles in the \abc\ light
curve are not matched by the PDD models, which exhibit more smooth variations.
Furthermore, while the focus of this study is the early evolution of \abc, the
PDD models evolve more rapidly to the red post-peak than the observations.
Nevertheless, the PDD models presented in \citet{2014MNRAS.441..532D} provide
several attractive explanations for the unusual features in the early behavior
of \abc, and small adjustments to the model (e.g., additional mixing of the
radioactive material, which \citeauthor{2014MNRAS.441..532D} only explore for
the DD model) may better match \abc.}

\subsection{Strong $^{56}Ni$ Mixing in the SN Ejecta}
\label{sec:Ni_mixing}

Having examined other possibilities we now consider whether the \textbf{early evolution} of \abc\ can be explained simply by invoking strong mixing in the 
SN ejecta. Strong mixing leads to \textbf{a faster initial rise as well as a more rapid evolution towards blue colors}.

\begin{figure}[]
  \centering
  \includegraphics[width=0.48\textwidth]{compare_Piro.pdf}
  \caption{  
  \textbf{Same as Figure~\ref{fig:dessart}, featuring models from
  \citet{2016ApJ...826...96P}. The amount of $^{56}$Ni mixing in the SN ejecta
  increases from the short-dash, light lines to the long-dash, dark lines.
  Unlike Figure~\ref{fig:dessart}, each model features the same
  $M_\mathrm{Ni}$, while the model names reflect the boxcar widths used to
  approximate the effects of mixing in the ejecta (see
  \citealt{2016ApJ...826...96P}). The top panel inset shows the
  residuals relative to the $0.25\,M_\odot$ model. The observations are best
  matched by the $0.25\,M_\odot$ model, i.e. the model with the most
  significant mixing.}
  }
  \label{fig:piro}
\end{figure}

Figure~\ref{fig:piro} compares the models from \citet{2016ApJ...826...96P} to
\abc. The \citeauthor{2016ApJ...826...96P} models employ a piston driven
explosion to explode a single WD progenitor model. As the piston explosion
does not result in any nucleosynthesis, the distribution of $^{56}$Ni in the
ejecta must be prescribed by hand, which enables a study of the effects of
mixing on the resulting SN emission. Each model employs a fixed $0.5\,\sm$ of
$^{56}$Ni that has been distributed throughout the ejecta via boxcar averaging
(see their Figure~1). The resulting light curves are synthesized using the
SuperNova Explosion Code (\texttt{SNEC}; \citealt{2015ApJ...814...63M}), as
shown in Figure~\ref{fig:piro}. Broadly speaking, the results can be
summarized as follows: SNe with strong mixing exhibit a rapid rise and quickly
develop blue colors, whereas models where the $^{56}$Ni is confined to the
innermost layers of the ejecta remain faint for days after explosion and
\textbf{feature a gradual color evolution from the red to the blue. The model
with the strongest mixing (dark, long-dashed line in Figure~\ref{fig:piro})
best matches the observations of \abc. This is the only model we have found
that exhibits a flat $(B-V)_0$ color evolution in the days after explosion,
however, in detail this model is too red relative to the observations of \abc.
}

While the models from \citet{2016ApJ...826...96P} provide a good match to the
optical photometric evolution of \abc, they consistently overpredict the flux
in the UV. \textbf{They} also overpredict the
photospheric velocity of \abc\ by $\sim$2--3,000$\; \mathrm{km \; s}^{-1}$.
Future modeling efforts may improve the match between the simulations and
observations for \abc.

Both the PDD models and ejecta-mixing models show discrepancies with some
early observations of \abc. Nevertheless, we conclude that one, or both, of
these scenarios, which feature qualitatively similar predictions, is the most
likely explanation for \abc. Indeed, it may be the case that the typical
sequence of photometric and spectroscopic observations of young SNe Ia can
never distinguish between these two possibilities \citep{2017MNRAS.472.2787N}.

\section{\textbf{Discussion: The Emerging Sample of Young \sneia}}

The proliferation of high-cadence, time-domain surveys has led to several SNe Ia being discovered within $\lesssim$2\,d of first light in the past $\sim$decade. Observations probing the early evolution of these SNe allows us to place unique constraints on their progenitor systems and the corresponding explosion physics. This has revealed diversity in the earliest epochs after explosion and that commonly used SNe Ia templates do not match observations (e.g., \citealt{2012ApJ...744...38F}).

In \S\ref{sec:first_light} we compared our early observations of \abc\ to SN\,2011fe, which has the most comprehensive observations of the young SN Ia sample. Given its normal spectroscopic and photometric evolution, SN\,2011fe has been adopted as a standard for the early evolution of SNe Ia in many studies. While a detailed quantitative analysis is beyond the scope of this study, a qualitative examination of very young SNe Ia that otherwise exhibit normal spectra and evolution at and post-peak\footnote{This definition excludes iPTF\,14atg, which was shown to be subluminous with SN\,2002es-like spectra \citep{2015Natur.521..328C}.} reveals considerable diversity. In other words, at early times SN\,2011fe may not be the norm.

For SN\,2011fe the initial rise is well described by a $t^2$ power law, the $(B - V)_0$ colors evolve from the red to the blue in the $\sim$week after explosion, and the \ion{C}{2} present in the initial spectra is weak \citep{2011Natur.480..344N,2016ApJ...820...67Z,2012ApJ...752L..26P}. In contrast, \abc\ exhibits a near linear rise in flux, the $(B - V)_0$ colors are blue and roughly constant, and the \ion{C}{2} absorption is strong. Examining just these 3 qualitative features, SN\,2009ig is well matched to SN\,2011fe \citep{2012ApJ...744...38F}, while SN\,2013dy \citep{2013ApJ...778L..15Z}, SN\,2017cbv \citep{2017ApJ...845L..11H}, and \abc\ all bear a striking resemblance. SN\,2012cg, on the other hand, is intermediate to these two groups with weak C II and a relatively shallow early rise, like SN\,2011fe, but blue $(B - V)_0$ colors, like \abc\ \citep{2012ApJ...756L...7S,2016ApJ...820...92M}. SN\,2014J is intermediate in the other direction in that it exhibits a near linear rise \citep{2014ApJ...783L..24Z,2015ApJ...799..106G}, but the color evolution is very similar to SN\,2011fe \citep{2014ApJ...788L..21A}.\footnote{\ion{C}{2} is not detected in the spectra of SN\,2014J \citep{2014ApJ...784L..12G,2014ApJ...783L..24Z}, though the earliest spectra of SN\,2014J were obtained at a much later phase than the other SNe discussed here. \citet{2015ApJ...798...39M} find evidence for \ion{C}{1} in the NIR spectra of SN\,2014J, but this detection cannot constrain the relative strength of \ion{C}{2} and \ion{Si}{2} shortly after explosion.} That these early observations cannot be easily separated into two distinct groups suggests that it is unlikely that a single physical mechanism drives the diversity of SNe Ia at early times.

In the case of SN\,2012cg and SN\,2017cbv it has been argued that the early blue optical colors are indicative of interaction between the SN ejecta and a binary companion \citep{2016ApJ...820...92M,2017ApJ...845L..11H}. SN\,2017cbv is particularly remarkable in that the observations presented in \citet{2017ApJ...845L..11H} show a clearly resolved bump in the $U$, $B$, and $g'$ bands in the $\sim$5\,d after explosion. In \citeauthor{2017ApJ...845L..11H} it is found that the bump can be explained via the combination of ejecta-companion interaction and the normal evolution of a SN Ia. A challenge for this model, similar to \abc\ (see \S\ref{sec:companion}), is that it significantly overpredicts the UV brightness of the SN compared to what is observed. Indeed, in the case of SNe\,2012cg, 2017cbv, and \abc\ the $\mathrm{UV} - \mathrm{optical}$ colors are significantly redder than those observed in iPTF\,14atg. It is argued in \citet{2017ApJ...845L..11H} that several model assumptions, including: (i) ideal blackbody emission, (ii) a constant opacity, (iii) a simple power-law density profile for the ejecta, and (iv) spherical symmetry, may be incorrect, which could reconcile the discrepancy with the UV observations. Above, we argued for circumstellar interaction and strong $^{56}$Ni mixing as a possible explanation for \abc, and indeed \citet{2017ApJ...845L..11H} consider these possibilities for SN\,2017cbv as well. Separately, several arguments against companion interaction for SN\,2012cg are presented in \citet{2016arXiv161007601S}.

Ultimately, there are arguments in favor of and against each of the possibilities to explain the early emission from SNe Ia. Moving forward, more detailed models and simulations are needed to properly explain the observed diversity. No matter the correct explanation for the early behavior of SNe\,2013dy, 2017cbv, and \abc, the strong similarities between these events suggests that they may reflect a common physical origin.

\section{Conclusion}
\label{sec:conclusion}

We have presented observations of the extraordinarily early discovery of the 
normal Type Ia supernova \abc. Our fast-response follow-up 
campaign allowed us to draw the following conclusions:

\begin{itemize}
    \item Extrapolation of the early light curve shows that the initial detection of \abc\ occurred only $0.15\pm_{0.07}^{0.15}$~d after the time of first light, $t_0$.

    \item We find no evidence for detectable signatures of SN shock cooling or the collision of the SN ejecta with a non-degnerate binary companion.

    \item Assuming that $v_\mathrm{ph} \propto t^{-0.22}$, then $t_\mathrm{exp} \approx t_0$. A short dark phase, as this implies, is likely the result of either strong $^{56}$Ni mixing or interaction of the SN ejecta with diffuse material.

    \item The strong and short-lived carbon features seen in the 
earliest spectra of \abc\ can only be explained if there is incomplete burning. The pulsational delayed detonation models presented in \citet{2014MNRAS.441..532D} produce \ion{C}{2} absorption that is as strong as \ion{Si}{2} at very early phases.

    \item In contrast to SN\,2011fe, $(B - V)_0$ is $\sim$0.5 mag bluer for \abc\ $\sim$2\,d after $t_0$. Furthermore, the $(B - V)_0$ colors of \abc\ show no evolution over the first $\sim$7\,d  of observations.

    \item Finally, we show that the early light curve evolution and colors of \abc\ are well matched by the theoretical models of \citet{2016ApJ...826...96P} that include significant mixing of $^{56}$Ni into the outer layers of the SN ejecta.

\end{itemize}
%
Taken together, these observations all indicate that the nucleosynthetic products from the explosion are well mixed throughout the SN ejecta. There are elements of the PDD models from \citet{2014MNRAS.441..532D} that are attractive for explaining \abc, in particular, the strong \ion{C}{2} absorption seen at early times. In the future, it would be useful to investigate more detailed PDD models that incorporate strong $^{56}$Ni mixing to see if they better replicate the observations of \abc\ as it is otherwise difficult to distinguish between these two scenarios \citep{2017MNRAS.472.2787N}.\footnote{Both \citet{2014MNRAS.441..532D} and \citet{2016ApJ...826...96P} consider models with strong mixing \textit{and} interaction, but those studies do not replicate each of the salient features of \abc.}

Extremely early observations of young SNe provide a ``smoking
gun'' to probe the mixing level in the ejecta, which, in turn, is 
a result of the explosion mechanism. Wide-field, high-cadence surveys, such as the Zwicky Transient Facility \citep{2016PASP..128h4501B} and ATLAS \citep{2011PASP..123...58T,2013RSPTA.37120269T}, will discover a large number of very young SNe over the next few years, allowing us to extend our studies beyond single objects. While the sample of extremely young SNe Ia will grow by more than an order of magnitude, the detection of shock breakout cooling and ejecta-companion interaction will prove challenging. Given the diminutive size of WDs, the thermal emission following shock breakout can only be detected to $\sim 10\,\mathrm{Mpc}$ on 1-m class telescopes. Furthermore, only $\sim$10\% of single-degenerate progenitors are expected to give rise to detectable emission following the collision of the SN ejecta with the binary companion \citep{2010ApJ...708.1025K}. Despite these limitations, this study of \abc\ shows that the early detection of SNe Ia can probe explosion physics by measuring the amount of mixing in the SN ejecta. Moving forward, a large sample of such objects will enable strict constraints on the proposed explosion mechanisms for Type Ia SNe.

Finally, we close by emphasizing the importance of fast-response photometric and spectroscopic follow-up campaigns. Without the early recognition of the youth of this SN, and the associated follow-up, much of the analysis presented herein would not have been possible. The ability to trigger such observations is essential to improve our understanding of the physics of SNe Ia.

\acknowledgements

We are deeply grateful to R.~C.~Thomas for indulging a slew of questions regarding \ion{C}{2} in SNe. Similarly, it is our pleasure to buy a beer for R.~Amanullah and U.~Feindt for discussions regarding SALT2.

This paper utilizes the LCO infrastructure for rapid and regular monitoring of SNe, and we thank S.~Valenti and I.~Arcavi for their development efforts.

Much of the analysis presented herein would not have been possible without the help of several observers. We thank M.~West for taking the first spectrum of \abc\ as a ToO on the DCT. We also thank P.~GuhaThakurta, E.~C.~Cunningham, K.~A.~Plant, H.~Jang, and J.~Torres for executing a Keck ToO as part of the UC/Caltech partnership, and also the Gemini service observers for executing our ToO observations. Additionally, J.~Cohen, N.~Suzuki, V.~Ravi, R.~Walters, A.~Ho, H.~Vedanthamand, K.~De, and L.~Yan helped obtain data for this paper.

AAM is funded by the Large Synoptic Survey Telescope Corporation in support of the Data Science Fellowship Program. YC thanks support from a postdoc fellowship in the eScience institute, University of Washington.

DAH, CM, and GH are supported by NSF-1313484.

The Intermediate Palomar Transient Factory project is a scientific collaboration among the California Institute of Technology, Los Alamos National Laboratory, the University of Wisconsin, Milwaukee, the Oskar Klein Center, the Weizmann Institute of Science, the TANGO Program of the University System of Taiwan, and the Kavli Institute for the Physics and Mathematics of the Universe. This work was supported by the GROWTH project funded by the National Science Foundation under Grant No 1545949. Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the NASA. This work makes use of observations from the LCO network. These results made use of the Discovery Channel Telescope at Lowell Observatory. Lowell is a private, non-profit institution dedicated to astrophysical research and public appreciation of astronomy and operates the DCT in partnership with Boston University, the University of Maryland, the University of Toledo, Northern Arizona University and Yale University.  The upgrade of the DeVeny optical spectrograph has been funded by a generous grant from John and Ginger Giovale. Based on observations made with the Nordic Optical Telescope, operated by the Nordic Optical Telescope Scientific Association at the Observatorio del Roque de los Muchachos, La Palma, Spain, of the Instituto de Astrofisica de Canarias.

\facility{DCT, Gemini:Gillett, Hale, Keck:I, Keck:II, LCOGT, PO:1.2\,m, PO:1.5\,m, VLT, Swift, OANSPM:HJT}

\software{\texttt{PTFIDE} \citep{2017PASP..129a4002M}, \texttt{FPipe}, \texttt{SALT2}, \texttt{SNID}, \texttt{lcogtsnpipe}, \texttt{UVOTSOURCE}}.

\appendix

\section{Photometric Light Curves}

The full photometric light curves of \abc\ are shown in Figure~\ref{fig:ugly_LC}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{lightcurve.pdf}
  \caption{UV, optical, and NIR light curves for \abc. 
  \textit{Top}: $BVgri$ light curves from the P48, P60, and LCO-1m 
  telescopes. The solid lines represent the best-fit model from 
  SALT2 (see \ref{sec:classification}).
  \textit{Bottom}: UV and NIR light curves from the \textit{Swift} 
  and SPM-1.5m telescopes, respectively.
  For both panels each light curve is represented with a different 
  color, while the different symbols correspond to different 
  instruments, as detailed in the legends. The legends also list 
  offsets applied to each light curve. Photometry is shown in the 
  AB mag system, with the exception of the $BV$ bands, which are 
  shown in the Vega system. The photometry shown has not been 
  corrected for line-of-sight extinction.
  }
  \label{fig:ugly_LC}
\end{figure}

\bibliographystyle{aasjournal}
\bibliography{ref}

\end{document}